### 2017-02-02 paper.tex
Wording of section 2

\documentclass[english,listof=totoc]{scrartcl}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{url}
\usepackage{amsmath}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage[authoryear]{natbib}
\usepackage[unicode=true,pdfusetitle,
 bookmarks=true,bookmarksnumbered=false,bookmarksopen=false,
 breaklinks=false,pdfborder={0 0 0},pdfborderstyle={},backref=false,colorlinks=false]
 {hyperref}\usepackage{mathtools}

\title{ML Estimation of Matrix--$t$ Distributions with Expectation Maximization}

\author{Zahari Kassabov\thanks{TIF-UNIMI-2016-5}\\
        Dipartimento di Fisica, Universit\`a di Torino and INFN, Sezione di Torino\\
		TIF Lab, Dipartimento di Fisica, Universit\`a di Milano\\
        \texttt{kassabov@to.infn.it}\\
        \and
        Hsien-Ching Kao\\
        Wolfram Research, Inc.\\
		Champaign, Illinois 61820, USA\\
		\texttt{sp000088@gmail.com}
		}

\begin{document}

\maketitle

\begin{abstract}
We provide an expectation-maximization (EM) algorithm for the maximum
likelihood estimation of matrix--$t$ distributions and discuss its
practical implementation. The result can be applied to problems where
the parameters of a matrix--$t$ distribution need to be obtained from
data, such as the design of recommendation systems and remote sensing
algorithms.
\end{abstract}

\section{Introduction}
The matrix--$t$ distribution is a generalization of the multivariate
$t$ distribution to matrix variate domains. Similar to other $t$
distributions, the matrix--$t$ distribution is a scale mixture of
a particular class of multivariate normal models: In particular, it is
a mixture of a matrix normal distribution, with a matrix-valued scale
parameter \citep{gupta1999matrix}.  Because of its mixture nature, the
matrix--$t$ distribution can be applied to building predictive models
with applications in tasks such as  spatial interpolation (for
example, to predict the concentration of pollution
\citep{KIBRIA2006785}) and recommendation systems
\citep{NIPS2007_3203}. Matrix models can be advantageous compared to
more traditional PCA methods because the correlation structure among
different observations can be captured in the scale matrix parameters,
and hence do not require the $i.i.d$ assumption
\citep{2013arXiv1309.6609G}. In such inferential problems, it is
required to obtain the parameters of the underlying model, given a set
of observed data samples. For matrix--$t$ distributions, the
parameters consist of a degrees of freedom parameter, a matrix
location parameter, and two scale matrix parameters.

The method of maximum likelihood (ML) is a common approach to estimate
the parameters of the distribution. However, for matrix--$t$
distributions, optimizing the likelihood directly is impractical since
the gradient equations are unwieldy. Instead, we can exploit the fact
that general for $t$ distributions, it is well-known that one can use
the mixture property and apply the expectation maximization (EM)
technique to optimize the likelihood function
\citep{10.2307/24305551}. EM can also be  applied to the particular
case of a matrix variate domains, by viewing the matrix--$t$
distribution as a mixture.

The related problem of determining a sample from a matrix--$t$
distribution given some observed entries of the matrix and parameters
of the model was studied in \cite{NIPS2007_3203}.

This paper is organized as follows: In section \ref{sec:matrixt} we
give a brief review on matrix--$t$ distributions, fix the notation and
give some matrix calculus results. In section \ref{sec:EM}, we derive
the EM formulae and provide relevant implementation details.  A brief
discussion is given in section \ref{sec:discussion}.

\section{Matrix--$t$ distributions}\label{sec:matrixt}

All matrices referred in this paper are real. The matrix determinant
is denoted as $|\cdot|$. We use the symbol $\mathbf{S}_{++}^p$ to
indicate the set of symmetric positive definite (SPD) matrices of
dimension $p$. The multivariate gamma function $\Gamma_{p}(\cdot)$,
given in terms of the standard gamma function $\Gamma(\cdot)$, is
%
\begin{equation}
\Gamma_{p}(x)=\pi^{p(p-1)/4}\prod_{j=1}^{p}\Gamma\left(x+\frac{1-j}{2}\right)
\ .\label{eq:multgammadef}
\end{equation}
%
We also define the multivariate digamma function
%
\begin{equation}
\psi_{p}(x):=\frac{d}{dx}\log\Gamma_{p}(x)=%
\sum_{j=1}^{p}\psi\left(x+\frac{1-j}{2}\right)\ .
\end{equation}

A $p
\times m$ random matrix $T$ is said to be matrix--$t$ distributed if
given the scale parameter $S$, a SPD matrix of dimensions $p\times p$,
the matrix $T$ is matrix normal distributed $N_{p,m}(M,S,\Omega)$ and
$S$ is itself inverse Wishart distributed $W^{-1}(\Sigma,\nu +p -1)$:
%
\begin{equation}
T\,|\,M,\Sigma,\Omega,\nu,S \sim N_{p,m}(M,S,\Omega) \ ,\label{eq:tdef1}
\end{equation}
and
\begin{equation}
S\,|\,\Sigma,\nu \sim W^{-1}(\Sigma,\nu+p-1)\ .\label{eq:tdef2}
\end{equation}
%
Therefore, the matrix--$t$ distribution $T_{p,m}(M,\Sigma,\Omega,\nu)$
of dimensions $p\times m$ is specified by 4 parameters: the location
matrix $M$ of dimensions $p\times m$ , the row and column SPD scale
matrices $\Sigma$ and $\Omega$ of dimensions $p\times p$ and $m\times
m$, respectively, and the degrees of freedom parameter $\nu>0$.

The density of the matrix normal distribution $N_{p,m}(M,S,\Omega)$
\begin{equation}
\frac{1}{(2\pi)^{pm/2}|\Omega|^{p/2}|S|^{m/2}}%
\exp\left(-\frac{1}{2}\textrm{tr}\left[\Omega^{-1}(X-M)^{\textrm{t}}S^{-1}(X-M)\right]\right)
\label{eq:ndistpdf}
\end{equation}
at $X$,  while the density of the inverse Wishart distribution
$W^{-1}(\Sigma,n)$ with $n>p-1$ is
\begin{equation}
\frac{|\Sigma|^{\frac{n}{2}}}{2^{\frac{n p}{2}}%
|S|^{\frac{n+p+1}{2}}\Gamma_{p}(\frac{n}{2})}%
\exp\left(-\frac{1}{2}\textrm{tr}\left[\Sigma S^{-1}\right]\right)
\label{eq:wdistpdf}
\end{equation}
at $S$. Here $\Gamma_{p}(\cdot)$ is the multivariate gamma function
defined in Eq.~\eqref{eq:multgammadef}.

From the definitions in Eqs.~\eqref{eq:tdef1} and \eqref{eq:tdef2},
and the densities given above, it follows that the density function
$f(T;\left\{M,\Sigma,\Omega,\nu\right\})$ of a matrix--$t$
distribution $T_{p,m}(M,\Sigma,\Omega,\nu)$ satisfies
\begin{equation}
\frac{\Gamma_{p}\left(\frac{p+\nu+m-1}{2}\right)}%
{\Gamma_{p}\left(\frac{\nu+p-1}{2}\right)%
\pi^{\frac{mp}{2}}|\Omega|^{\frac{p}{2}}|\Sigma|^{\frac{m}{2}}}%
|I_p+\Sigma^{-1}(T-M)\Omega^{-1}(T-M)^{\textrm{t}}|^{\frac{1-m-p-\nu}{2}}
\label{eq:matrixtpdf}
\end{equation}
%
with $I_p$ the identity matrix of dimension $p$. The marginalization
over $S$ can be obtained by using the Jacobain determinant of an
inverse SPD matrix $U=S^{-1}$
%
\begin{equation}
\left|\frac{\partial U}{\partial S}\right|=|S|^{-p-1}
\end{equation}
and the integral identity \citep{gupta1999matrix}
\begin{equation}
\int_{\mathbf{S}_{++}^p}e^{-\textrm{tr}[S A]}|S|^{\alpha}\,dS=%
\Gamma_{p}\left(\alpha+\frac{1+p}{2}\right)|A|^{-\alpha-\frac{1+p}{2}}
\ .\label{eq:intmultgammadef}
\end{equation}
Where $A$ is a $p\times p$ nonsingular matrix.

Similar to matrix normal distributions, matrix--$t$ distributions are
defined up to a multiplicative constant. It is easy to see from the
density function \eqref{eq:matrixtpdf} that
$T_{p,m}(M,c\,\Omega,\Sigma/c,\nu)$ has the same distribution as
$T_{p,m}(M,\Omega,\Sigma,\nu)$ for any constant $c>0$. This property
needs to be taken into account in an iterative ML estimation algorithm
in order to be able to reach convergence.

\section{EM formulae for matrix--$t$ distribution}\label{sec:EM}

Given a random variable $X$ with probability density (or mass) function $p_{\theta}(\cdot)$. Assume the density (or mass) function is a marginalization over some unobserved random state $Z$, i.e., $p_{\theta}(x)=\mathbb{E}_{\theta}\left[p_{\theta}(x|Z)\right]$. With observations $\left\{x_i\right\}_{i\in I}$ the classical EM method for ML estimation consists of two steps:

\begin{itemize}
\item At the $k$-th iteration, compute the conditional expectations with respect to $Z$ for each observation $x_i$
\begin{equation}
Q_i(\theta,\theta_k)=\mathbb{E}_{\theta_k}\left[\log p_{\theta}(x_i,Z)|x_i\right].
\end{equation}

\item Update $\theta$ by maximizing $Q(\theta,\theta_k):=\sum_{i\in I}Q_i(\theta,\theta_k)$ with respect to $\theta$, i.e.,
\begin{equation}
\theta_{k+1} = \textrm{argmax}_{\theta}\,Q(\theta,\theta_k).
\end{equation}
\end{itemize}

\subsection{Expectation step}

Given observed matrix data $\left\{T_j\right\}_{j\in J}$ for each observation $T_j$ the E-step of the matrix--$t$ distribution entails computing the following integral at step $k$
\begin{equation}
Q_j(\theta,\theta_k)=f(T_j;\theta_k)^{-1}\int_{\mathbf{S}_{++}^p}f(T_{j},S;\theta_k)\log f\left(T_{j},S;\theta\right)\,dS,\label{eq:expdef}
\end{equation}
where $\theta=\left\{M,\Sigma,\Omega,\nu\right\}$, $\theta_k=\left\{M_k,\Sigma_k,\Omega_k,\nu_k\right\}$, and $f(T,S;\theta)$ is the full likelihood of the distribution with the hidden scale parameter $S$.

We can split the integral in Eq.~\eqref{eq:expdef} by the terms of $\log f(T_{j},S;\theta)$. Eliminating a global factor $\frac{1}{2}$ and constant terms which do not depend on the parameters in $\theta$ we have
\begin{equation}
\begin{split}\log f(T_{j},S;\theta)\sim -\nu\log|S|-\textrm{tr}\left[S^{-1}\left\{\Sigma+(T_j-M)\Omega^{-1}(T_j-M)^{\textrm{t}}\right\}\right]\\
-\nu p\log2-p\log |\Omega|+(\nu+p-1)\log |\Sigma|-2\log\Gamma_{p}\left(\frac{\nu+p-1}{2}\right).
\end{split}\label{eq:logf}
\end{equation}
The expectation computation requires computing the following two integrals:
\begin{equation}
\int_{\mathbf{S}_{++}^p}f(T_j,S;\theta_k)\log |S|\,dS, \qquad 
\int_{\mathbf{S}_{++}^p}f(T_j,S;\theta_k) S^{-1}\,dS.
\end{equation}
It can be done by using equalities derived by taking the derivative of Eq.~\eqref{eq:intmultgammadef} with respect to $\alpha$ and matrix $A$, respectively, and this gives
\begin{equation}
\int_{\mathbf{S}_{++}^p}e^{-\textrm{tr}\left[SA\right]}|S|^{\alpha}\log |S|\,dS=\Gamma_{p}\left(\alpha+\frac{1+p}{2}\right)|A|^{-\alpha-\frac{1+p}{2}}\left\{\gamma_{p}\left(\alpha+\frac{1+p}{2}\right)-\log |A|\right\},\label{eq:intderivative1}
\end{equation}
\begin{equation}
\int_{\mathbf{S}_{++}^p}e^{-\textrm{tr}\left[SA\right]}|S|^{\alpha}S\,dS=\left(\alpha+\frac{1+p}{2}\right)\Gamma_{p}\left(\alpha+\frac{1+p}{2}\right)|A|^{-\alpha-\frac{1+p}{2}}A^{-1}.\label{eq:intderivative2}
\end{equation}

Utilizing the equalities \eqref{eq:intderivative1} and \eqref{eq:intderivative2} with adequate rearrangements we obtain the closed form expression of $Q_j(\theta,\theta_k)$ while ignoring the irrelevant constant terms
\begin{equation}
\begin{split}Q_j(\theta,\theta_k) \sim \nu\left\{\log|W_j^{(k)}|+\psi_{p}\left(\frac{m+\nu_k+p-1}{2}\right)\right\}-2\log\Gamma_{p}\left(\frac{\nu+p-1}{2}\right)-p\log|\Omega|\\
-(m+\nu_k+p-1)\textrm{tr}\left[W_j^{(k)}\left\{\Sigma+(T_j-M)\Omega^{-1}(T_j-M)^{\textrm{t}}\right\}\right]+(\nu+p-1)\log |\Sigma|.
\end{split}
\label{eq:fsimp}
\end{equation}
Here the matrices $\left\{W_j^{(k)}\right\}_{j\in J}$ are defined as
\begin{equation}
W_j^{(k)}:=\left\{\Sigma_k+(T_j-M_k)\Omega_k^{-1}(T_j-M_k)^{\textrm{t}}\right\}^{-1}.
\end{equation}

As $\Sigma_k$ and $\Omega_k$ are both SPD, each matrix $W_j^{(k)}$ is also SPD. The computation of $W_j^{(k)}$ can be done efficiently using Cholesky factorization via LAPACK routines and the BLAS routines for symmetric matrix update. When $p$ is much greater than $m$, i.e., the dimension of $\Sigma$ is much larger than $\Omega$, one can apply Sherman--Morrison--Woodbury formulae \citep{doi:10.1137/1023004} to reduce the cost of inverting the $p\times p$ matrix for each $W_j^{(k)}$. By using the formulae $W_j^{(k)}$ can be rewritten in the following equivalent form
\begin{equation}
W_j^{(k)}=\Sigma_k^{-1}-\Sigma_k^{-1}U_j\left(I_m+U_j^{\textrm{t}}\Sigma_k^{-1}U_j\right)^{-1}U_j^{\textrm{t}}\Sigma_k^{-1}\label{eq:SMW}
\end{equation}
with $U_j=(T_j-M_k)\Omega_k^{-1/2}$ and $\Omega_k^{1/2}$ is the upper Cholesky factor of $\Omega_k$. Note that in Eq.~\eqref{eq:SMW} one inverts a $m\times m$ matrix for each $j$ and the inversion (or factorization) of $\Sigma_k$ only needs to be done once in each iteration.

\subsection{Maximization step}

To maximize $Q(\theta,\theta_k)$ with respect to $\theta$ we compute the gradient of each parameter and equate them to $0$ to obtain $\theta_{k+1}$. For parameter $M$ we have
\begin{equation}
\sum_{j\in J}W_j^{(k)}\left(T_j-M\right)\Omega^{-1}=0.
\end{equation}
As $\Omega$ is nonsingular we obtain $M_{k+1}$ as an average of the input data weighted by $W_j^{(k)}$
\begin{equation}
M_{k+1}=\left(\sum_{j\in J}W_j^{(k)}\right)^{-1}\sum_{j\in J}W_j^{(k)}T_j.\label{eq:M}
\end{equation}
Maximizing $Q(\theta,\theta_k)$ with respect to $\Omega$ we obtain $\Omega_{k+1}$
\begin{equation}
\Omega_{k+1}=\frac{m+\nu_k+p-1}{p}\left(\frac{1}{|J|}\sum_{j\in J}(T_j-M_{k+1})^{\textrm{t}}W_j^{(k)}(T_j-M_{k+1})\right).\label{eq:Omega}
\end{equation}
For the rest of the parameter(s), we consider two possible situations.
\begin{itemize}
\item $\Sigma$ and $\nu$ are unknown: In this situation, maximizing $Q(\theta,\theta_k)$ with respect to $\Sigma$ we obtain $\Sigma_{k+1}$ in terms of the inverse of the mean of $W_j^{{(k)}}$
\begin{equation}
\Sigma_{k+1}(\nu)=\frac{\nu+p-1}{m+\nu_k+p-1}\left(\bar{W}^{(k)}\right)^{-1},\quad \bar{W}^{(k)}:=\frac{1}{|J|}\sum_{j\in J}W_j^{(k)}.\label{eq:Sigma}
\end{equation}
$\Sigma_{k+1}$ depends linearly on $\nu$. Due to this, we substitute $\Sigma_{k+1}$ back into $Q(\theta,\theta_k)$ and maximize it with repect to $\nu$. We remark that
\[\tilde{Q}(\nu):=Q\left(\left\{M_{k+1},\Sigma_{k+1}(\nu),\Omega_{k+1},\nu\right\},\theta_k\right)\]
is a concave function of $\nu$ for all $\nu>0$. The maximum only attains at positive $\nu$ if $\tilde{Q}'(0)$ is positive. In this situation we obtain $\nu_{k+1}$ by solving the equation $\tilde{Q}'(\nu)=0$ which has the form
\begin{equation}
h(\nu+p-1)-h(m+\nu_k+p-1)+\frac{1}{|J|}\sum_{j\in J}\log|W_j^{(k)}|-\log |\bar{W}^{(k)}|=0\label{eq:nu1}
\end{equation}
with $h(x):=p\log(x)-\psi_p(x/2)$.
\item $\Sigma$ is known and $\nu$ is unknown: In this situation, $Q\left(\left\{M_{k+1},\Sigma,\Omega_{k+1},\nu\right\},\theta_k\right)$ is also a concave function of $\nu$. The gradient equation of $\nu$ is given by
\begin{equation}
\log|\Sigma|-\log |\bar{W}^{(k)}|+\psi_{p}\left(\frac{m+\nu_k+p-1}{2}\right)-\psi_{p}\left(\frac{\nu+p-1}{2}\right)=0.\label{eq:nu2}
\end{equation}
\end{itemize}

When $\Omega$ and $\Sigma$ are both unknown, the updating steps given in Eqs.~\eqref{eq:Omega} and \eqref{eq:Sigma} increases the likelihood of matrix--$t$ distribution monotonically but the parameters $\Omega_k$ and $\Sigma_k$ may not converge due to the non-identifiability of matrix--$t$ mentioned at the end of section \ref{sec:matrixt}. This issue has also been discussed in \citep{2013arXiv1309.6609G} for ML estimation of matrix normal distributions. However, instead of fixing a single component in one of the scale matrices, we suggest setting the mean trace ratio
\begin{equation}
\frac{\textrm{tr}[\Omega_k]/m}{\textrm{tr}[\Sigma_k]/p}
\end{equation}
of $\Omega_k$ and $\Sigma_k$ to 1 by proper scaling at the end of each M-step. This ensures the convergence of the algorithm and also avoids possible numerical issues due to the possible large scale separation between $\Omega_k$ and $\Sigma_k$ during the iterations.

\subsection{Parameters initialization}

For EM method, the initial values for the parameters has significant effect on the convergence behavior as the algorithm does not guarantee reaching the global optimum. While this is less of an issue for matrix--$t$ distributions as the ML estimation is a convex optimization problem in parameters $\left\{M,\Sigma^{-1},\Omega^{-1},\nu\right\}$, it is still desirable to have an efficient method for selecting good initial values to reduce the number of EM iterations for reaching the required accuracy as the classical log EM is a first order algorithm.

For the location parameter $M$, the component-wise medians and spatial median \citep{Vardi15022000} with distance induced by Frobenius norm
\begin{equation}
d(T,T')=\sqrt{\textrm{tr}\left[(T-T')(T-T')^{\textrm{t}}\right]}
\end{equation}
are proper choices we found in the implementation. Both estimators can be computed with time complexity $O(Nmp)$ with $N$ the number of matrices in the given data.

For the rest of the parameters, we center the data by the initial values of $M$ and use the property of low dimensional projection of matrix--$t$ distributions to obtain the initial values. It is known that if $T\sim T_{p,m}(\mathbf{0},\Sigma,\Omega,\nu)$, then $Tv$ is multivariate $t$ distributed with scale parameter $\left(\frac{v^{\textrm{t}}\Omega v}{\nu}\right)\Sigma$ and the same degress of freedom parameter $\nu$ for any nonzero $m$ dimensional vector $v$ \citep{gupta1999matrix}. This allows using efficient routines for estimating multivariate $t$ distributions to retrieve the initial values.

\section{Discussion}\label{sec:discussion}

In this paper we derive and provide the EM formulae for ML estimation of matrix--$t$ distributions. The EM algorithm here mainly consists of computing the weight matrices $W_j^{(k)}$ and we provide some guidance on its numerical implementation. A method of choosing the initial values for the parameters in EM is briefly discussed.

In principle, the computation burden of estimating matrix--$t$ distributions can be reduced significantly with the EM algorithm discussed here. However, due to the lack of optimized routines which dedicate to linear algebra computations of matrix-valued vector components, the algorithm described here can be difficult to optimize in the actual implementation. Also, scalability with large data can be an issue as the computation involves dense matrix inversion for each observation during the EM iterations. It is possible to alleviate the cost by considering matrix--$t$ distributions with structured scale matrix parameters which allow fast factorization. The EM relations described here can still be applied but with some modifications in the M-step. The application of structured scale matrix parameters in matrix--$t$ inference will be future work.

\section*{Acknowledgement}
This work is not supported by any grant and has not been presented in any regional or international meetings.

\bibliographystyle{unsrtnat}
\bibliography{references}
\end{document}